---
title: "Descriptive stats: vowels and fricatives"
output: html_document
date: '2022-03-27'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = F,
                      message = F)
```

## data loading

```{r loaddata}
# Packages
library(tidyverse)
library(Hmisc)
library(DescTools)
library(stats)
library(tibble)
library(ggpubr)

# Load data as tibble
# Grand average data:
# For analyzing questionnaires descriptive stats
# And correlation matrix
data_ave = read_delim("gender_grand_ave.txt", delim = '\t')
# randomly show 4 rows in the df
sample_n(data_ave, 4)

# count the number of men and women
# it is for checking whether the coding for men and women are correty
data_ave %>% count(data_ave$sex)

# Vowel and fricative df
# For box plot series
# that explores effects of linguistic factors

# vowel
df_v = read_delim("gd_vowels_recode.txt", delim = '\t')
head(df_v)
# Delete unncessary columns
df_v <- select(df_v, -c(AQ1:BAP_rigid))

# Fricative
df_f = read_delim("gd_fricatives_recode.txt", delim = '\t')
head(df_f)

# Also delete the unnecessary columns
df_f <- select(df_f, -c(AQ1:BAP_rigid))

```

## Preliminary analysis on questionnaires

Use data averaged across items for descriptive stats analysis on the questionnaires and the acoustic analysis.

- Descriptive stats of questionnaires

```{r averageDF, include = TRUE, warning = FALSE}
head(data_ave)
# change the head name of GEPAQ_F column
data_ave <- rename(data_ave, EPAQ_F = GEPAQ_F)
data_ave

#descriptive stats for questionnaires
#Table 1
data_ave %>% select(speaker, sex, TMF, EPAQ_F) %>% 
  gather(questionnaire, scores, TMF:EPAQ_F) %>%
  group_by(questionnaire, sex) %>% 
  summarise(M = mean(scores), SD = sd(scores), 
            MIN = min(scores), MAX = max(scores)) %>% 
  t()

```

- Use t-test to examine whether male and female ratings on their femininity-masculinity levels differ significantly from each other

``` {r}
# t-test for the two questionnaires

t.test(TMF ~ sex, data = data_ave)

t.test(EPAQ_F ~ sex, data = data_ave)

```


- box plots and regression lines of the two questionnaires

```{r plotsofquestionnaires, include = TRUE, warning = FALSE}
# box plots
# require ggpubr library
# for showing the t-test results in the box plots

EPAQ_plot <- 
  ggplot(data_ave, aes(x = sex, y = EPAQ_F, color = sex)) +
  geom_boxplot()+
  stat_compare_means(
    method = "t.test",
    aes(label = ..p.signif..), 
    label.x = 1.5, #specify position of significant marking
    label.y = 40) +
  theme_classic()+
  labs(y = "EPAQ-F+",
       x = "sex") + 
  scale_x_discrete(labels=c("f" = "female", 
                            "m" = "male"))+
  theme(legend.position = "none") #remove legend

TMF_plot <- ggplot(data_ave, aes(x = sex, y = TMF)) +
  geom_boxplot(aes(color=sex))+
  stat_compare_means(
    method = "t.test", 
    aes(label = ..p.signif..),
    label.x = 1.5, 
    label.y = 40)+
  theme_classic() +
  labs(y = "TMF",
       x = "sex") + 
  scale_x_discrete(labels=c("f" = "female", 
                            "m" = "male"))+
  theme(legend.position = "none") 

##plot the regression line for two questionnaires with R square and p value
reg_plot <- ggplot(data_ave, aes(x = TMF, y = EPAQ_F, color = sex)) +
  geom_point() + # scatter plot
  geom_smooth(method = "lm", # add a linear regression line
              formula = y ~ x, # the formular for the linear regression
              se = TRUE) + # add confidence interval 
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
           label.x = 3, 
           label.y.npc = c(0.2, 0.2), #adjust the label in y axis
           size = 4)+
  theme_classic() +
  labs(y = "EPAQ-F+",
       x = "TMF") + 
  theme(legend.position = "none")

#arrange three plots
ggarrange(reg_plot,
          EPAQ_plot,
          TMF_plot,
          ncol = 3, 
          nrow = 1,
          widths = c(2.5, 1.5, 1.5)) #adjust the size of each coloumn

```


# Box plots of fricative and vowel data

## fricatives
- Explore the effect of prosodic positions and syllable positions on the CoG values of /s/

```{r}
#code for figure 3
#refinement includes:
#changing the color for the group
#theme change into classic

#seperate the data into onset data and coda data
#seperate the data into onset data and coda data
data_onset = df_f[df_f$syllable_position=="a_onset",] #don't forget the , in the end!!!
data_coda = df_f[df_f$syllable_position=="b_coda",]

#box plot for cog_onset value for phrasal position and target filler order
onset_plot <- ggplot(data_onset, aes(x = phrase_position, 
                                     y = cog,
                                     color=filler_target_order)) +
  geom_boxplot() +  
  scale_color_manual(breaks = c("a_target_filler", 
                                "b_filler_target"),
                     labels = c("a_target_filler" = "target filler",
                                "b_filler_target" = "filler target"),
                     values= c("coral", "forestgreen"))+
  labs(title = "CoG of onset /s/",
       x = "Phrase position", 
       y = "CoG (Hz)",
       color  = "utterance positions")+
  scale_x_discrete(labels=c("a_initial" = "initial", 
                            "b_final" = "final"))+
  scale_y_continuous(limit  = c(3000, 11000),
                     labels = scales::comma_format()
                     ) +
  theme_classic()+
  theme(plot.title = element_text(face = "bold", 
                                  size = (12),
                                  hjust = 0.5), #center the plot title
        axis.title = element_text(size = (9)),
        legend.title = element_text(size = (10)))

#box plot for cog_coda value for phrasal position and target filler order
coda_plot <- ggplot(data_coda, aes(x = phrase_position, 
                                     y = cog,
                                     color=filler_target_order)) +
  geom_boxplot() +  
  #change the legend content
  scale_color_manual(breaks = c("a_target_filler", 
                                "b_filler_target"),
                     labels = c("a_target_filler" = "target filler",
                                "b_filler_target" = "filler target"),
                     values= c("coral", "forestgreen"))+
  labs(title = "CoG of coda /s/",
       x = "Phrase position", 
       y = "CoG (Hz)",
       color  = "utterance positions")+
  scale_x_discrete(labels=c("a_initial" = "initial", 
                            "b_final" = "final"))+
  scale_y_continuous(limit  = c(3000, 11000),
                     labels = scales::comma_format()
  ) +
  theme_classic()+
  theme(plot.title = element_text(face = "bold", 
                                  size = (12),
                                  hjust = 0.5), #center the plot title
        axis.title = element_text(size = (9)),
        legend.title = element_text(size = (10)))

#combine thes two ggplots together
#require library ggpubr
ggarrange(onset_plot,coda_plot,
          ncol = 2, nrow = 1,
          common.legend = TRUE, legend = "bottom") #common legend


```


## vowels

- explore the effect of prosodic positions, consonantal contexts and hyper-articulation conditions on vowel space size

```{r}

#updated Figure 4 boxplot for vowels
#phrase positions for vowel data
phrase_v_plot <- ggplot(df_v, aes(x = filler_target_order, 
                                           y = v_space_bk,
                                           color=phrase_position)) +
  geom_boxplot() +  
  #change the legend content
  scale_color_manual(breaks = c("a_initial", 
                                "b_final"),
                     labels = c("a_initial" = "initial",
                                "b_final" = "final"),
                     values= c("coral", "forestgreen"))+
  labs(x = "Utterance position", 
       y = "Vowel space (bark^2)",
       color  = "Phrase positions")+
  scale_x_discrete(labels=c("a_target_filler" = "target filler", 
                            "b_filler_target" = "filler target"))+
  scale_y_continuous(limit  = c(0, 35)) +
  theme_classic()+
  theme(axis.title = element_text(size = (9)),
        legend.title = element_text(size = (9)),
        legend.position = c(0.75,0.85)) #c(0,0) bottom left; c(1,1) "top right"


#boxplot for consonant contexts
consonant_v_plot <- ggplot(df_v, aes(x = consonants, 
                                              y = v_space_bk,
                                              color = consonants)) +
  geom_boxplot() + 
  stat_compare_means(
    method = "t.test",
    aes(label = ..p.signif..), label.x = 1.5, label.y = 33
  )+ 
  scale_color_manual(values= c("blueviolet", "dodgerblue3"))+
  labs(x = "Consonant contexts", 
       y = "Vowel space (bark^2)")+
  scale_x_discrete(labels=c("bt" = "b_t", 
                            "hd" = "h_d"))+
  scale_y_continuous(limit  = c(0, 35)) +
  theme_classic()+
  theme(axis.title = element_text(size = (9)),
        legend.position = "none") 

#consonant_v_plot


#boxplot for hyper_contexts
hyper_v_plot <- ggplot(df_v, aes(x = hyper_context, 
                                          y = v_space_bk,
                                          color = hyper_context)) +
  geom_boxplot() + 
  stat_compare_means(
    method = "t.test",
    aes(label = ..p.signif..), label.x = 1.5, label.y = 33
  )+ 
  scale_color_manual(values= c("blueviolet", "dodgerblue3"))+
  labs(x = "Hyperarticulation", 
       y = "Vowel space (bark^2)")+
  scale_x_discrete(labels=c("a_no" = "no", 
                            "b_yes" = "yes"))+
  scale_y_continuous(limit  = c(0, 35)) +
  theme_classic()+
  theme(axis.title = element_text(size = (9)),
        legend.position = "none")

# Combine the plots together
ggarrange(phrase_v_plot,
          consonant_v_plot,
          hyper_v_plot,
          ncol = 3, 
          nrow = 1,
          align = "hv",
          widths = c(2, 1.5, 1.5)) #adjust the size of each coloumn

```

## Scatter-histogram plots of fricatives and vowels 

Pearson correlation analysis on acoustic signals and two questionnaires eliciting speakers' self-rated femininity-masculinity

### define the function for the plot

```{r}

# function for plotting scatter-histogram
scat_hist_plot <- function(df, 
                             x, # string for x in scatter plot
                             y, # string for y in scatter plot
                             xbins, # num of bins for x axis
                             ybins, # num of bins for y axis
                             x_title, # x-axis name
                             y_title, # y-axis name
                             x_limit, # x axis limits, a vector
                             y_limit, # y axis limits, a vector
                             x_break, # x axis breaks, a vector
                             y_break, # y axis breaks, a vector
                             results_position){ # position for listing R square
  # scatter plot
  scatter_plot <- ggplot(df, 
                         aes_string(x = sprintf("%s", x), 
                                    y = sprintf("%s", y),
                                    color = sprintf("%s", "sex")) 
                         #color = sex needed for stat_cor to list both R square
  ) +
    scale_x_continuous(limit  = x_limit,
                       breaks = x_break) +
    scale_y_continuous(limit  = y_limit,
                       breaks = y_break) +
    geom_point(aes(color = sex, shape = sex)) + # scatter plot
    geom_smooth(aes(color = sex, fill = sex),
                method = "lm", # add a linear regression line
                formula = y ~ x, # the formular for the linear regression
                se = TRUE # add confidence interval
    ) +  
    theme_classic() + 
    stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")),
             label.x = results_position) + 
    theme(legend.position = "none")+
    labs(x= sprintf("%s", x_title),
         y = sprintf("%s", y_title))
  # density plot for x
  xdensity <- ggplot(df, aes_string(x = sprintf("%s", x))) +
    scale_x_continuous(limit  = x_limit,
                       breaks = x_break) + 
    geom_histogram(aes(y = ..density.., color = sex, fill = sex),
                   alpha = 0.4,
                   position = "identity",
                   bins = sprintf("%d", xbins)
    ) +
    geom_density(aes(color = sex), size = 1)+
    labs(x = NULL, y = NULL)+
    theme_classic() +  
    theme(#panel.background = element_rect(fill = "white"), no use
          axis.text = element_blank(),
          legend.position = "none")
  # desity plot for y
  ydensity<- ggplot(df, aes_string(y = sprintf("%s", y))) + 
    scale_y_continuous(limit  = y_limit, #,
                       #labels = scales::comma_format()
                       breaks = y_break) + 
    geom_histogram(aes(x = ..density.., color = sex, fill = sex),
                   alpha = 0.4,
                   position = "identity",
                   bins = sprintf("%d", ybins)
    ) + 
    geom_density(aes(color = sex), size = 1) +
    labs(x = NULL, y = NULL)+
    theme_classic()+ 
    theme(axis.text = element_blank(),
          legend.position = "none")
  
  require("gridExtra")
  output_plot <- ggarrange(xdensity, 
                           NULL, 
                           scatter_plot, 
                           ydensity, 
                           ncol = 2, nrow = 2,  
                           align = "hv", 
                           widths = c(4, 1.5), heights = c(1.5, 4)
                           )
  return(output_plot)}

```

- load parameters
``` {r}
# predefined limits of x and y axis
# predefined breaks for histogram
TMF_limit <- c(0, 50)
TMF_breaks <- c(5, 10, 15, 20, 25, 30, 35, 40, 45)
GEPAQ_breaks <- c(14, 18, 22, 26, 30, 34, 38, 42, 46)
GEPAQ_limit <- c(10, 50)
COG_limit <- c(4500, 11000)
COG_breaks <- c(5000, 6000, 7000, 8000, 9000, 10000)
vspace_limit <- c(0,30)
vspace_breaks <- c(5, 10, 15, 20, 25)

```

### fricatives 

```{r}
tmf_plot_fric <- scat_hist_plot(data_ave,
                                    x = "TMF", 
                                    y = "COG_all",
                                    xbins = 16,
                                    ybins = 18,
                                    x_title = "TMF",
                                    y_title = "CoG (Hz)",
                                    x_limit = TMF_limit,
                                    y_limit = COG_limit,
                                    x_break = TMF_breaks,
                                    y_break = COG_breaks,
                                    results_position = 5)

geqaq_plot_fric <- scat_hist_plot(data_ave, 
                                      x = "EPAQ_F", 
                                      y = "COG_all",
                                      xbins = 20,
                                      ybins = 18,
                                      x_title = "EPAQ-F+",
                                      y_title = "CoG (Hz)",
                                      x_limit = GEPAQ_limit,
                                      y_limit = COG_limit,
                                      x_break = GEPAQ_breaks,
                                      y_break = COG_breaks,
                                      results_position = 15)

ggarrange(tmf_plot_fric,geqaq_plot_fric,
          ncol = 2, nrow = 1,
          align = "hv",
          widths = c(1, 1),
          heights = c(1,1))

```


### vowels
```{r}

tmf_plot_v <- scat_hist_plot(data_ave, 
                                 x = "TMF", 
                                 y = "v_space_bk",
                                 xbins = 16,
                                 ybins = 14,
                                 x_title = "TMF",
                                 y_title = "Vowel space (bark^2)",
                                 x_limit = TMF_limit,
                                 y_limit = vspace_limit,
                                 x_break = TMF_breaks,
                                 y_break = vspace_breaks,
                                 results_position = 5)

geqaq_plot_v <- scat_hist_plot(data_ave, 
                                   x = "EPAQ_F", 
                                   y = "v_space_bk",
                                   xbins = 20,
                                   ybins = 14,
                                   x_title = "EPAQ-F+",
                                   y_title = "Vowel space (bark^2)",
                                   x_limit = GEPAQ_limit,
                                   y_limit = vspace_limit,
                                   x_break = GEPAQ_breaks,
                                   y_break = vspace_breaks,
                                   results_position = 15)

ggarrange(tmf_plot_v,geqaq_plot_v,
          ncol = 2, nrow = 1,
          align = "hv",
          widths = c(1, 1),
          heights = c(1,1))

```

